% !TeX encoding = UTF-8
% !TeX program = pdflatex
% !TeX spellcheck = en_US
\documentclass[binding=0.6cm,LaM,oneside]{sapthesis} % LaM for a Laurea Magistrale
\usepackage{microtype}
\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}

\makeatletter
\def\BState{\State\hskip-\ALG@thistlm}
\makeatother

\usepackage{lineno}
\modulolinenumbers[5]
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{amssymb,amsmath,nccmath}
\usepackage{cclicenses}
\usepackage{makecell}
\usepackage{lscape,array}
\newcolumntype{C}[1]{>{\centering\arraybackslash}p{#1}} 
\usepackage[thin, , thinc]{esdiff}
\usepackage{subcaption}
\usepackage{caption}
\usepackage{framed}  
\usepackage{subcaption}
\usepackage[font=small,skip=0pt]{caption}

\PassOptionsToPackage{hyphens}{url}
\hypersetup{
	colorlinks = true, % Colours links instead of ugly boxes
	urlcolor = black, % Colour for external hyperlinks
	linkcolor = black, % Colour of internal links
	citecolor = black % Colour of citations
}

\usepackage{listings}
\usepackage{xcolor}
\usepackage{multirow}


\renewcommand\baselinestretch{1.5}
\colorlet{punct}{red!60!black}
\definecolor{background}{HTML}{EEEEEE}
\definecolor{delim}{RGB}{20,105,176}
\colorlet{numb}{magenta!60!black}

\lstdefinelanguage{json}{
	basicstyle=\normalfont\ttfamily,
	numbers=left,
	numberstyle=\scriptsize,
	stepnumber=1,
	numbersep=8pt,
	showstringspaces=false,
	breaklines=true,
	frame=lines,
	backgroundcolor=\color{background},
	literate=
	*{0}{{{\color{numb}0}}}{1}
	{1}{{{\color{numb}1}}}{1}
	{2}{{{\color{numb}2}}}{1}
	{3}{{{\color{numb}3}}}{1}
	{4}{{{\color{numb}4}}}{1}
	{5}{{{\color{numb}5}}}{1}
	{6}{{{\color{numb}6}}}{1}
	{7}{{{\color{numb}7}}}{1}
	{8}{{{\color{numb}8}}}{1}
	{9}{{{\color{numb}9}}}{1}
	{:}{{{\color{punct}{:}}}}{1}
	{,}{{{\color{punct}{,}}}}{1}
	{\{}{{{\color{delim}{\{}}}}{1}
	{\}}{{{\color{delim}{\}}}}}{1}
	{[}{{{\color{delim}{[}}}}{1}
	{]}{{{\color{delim}{]}}}}{1},
}

\hypersetup{pdftitle={Investigation of static features for improved APT malware identification},pdfauthor={Leonardo Sagratella}}
\title{Investigation of static features for improved APT malware identification}
\author{Leonardo Sagratella}
\IDnumber{1645347}
\course{CyberSecurity}
\courseorganizer{Facoltà di Ingegneria dell'informazione, informatica e statistica}
\AcademicYear{2019/2020}
\copyyear{2020}
\advisor{Prof. Riccardo Lazzeretti}
\coadvisor{Dr. Giuseppe Laurenza}
\authoremail{leonardosagratella@gmail.com}
\begin{document}
	\frontmatter
	\maketitle
	\dedication{
		Dedicato alla mia famiglia, mio padre, mia madre e mia sorella che mi sono sempre stati vicini e mi hanno sempre sostenuto in questo percorso. Ai miei compagni di corso conosciuti in questi anni, in particolare Davide, Gianluca e Gianmarco grazie davvero. Dedicato ai miei piu grandi amici di sempre del campeggio Antonio, Cecilia, Federico e Lucia, ma in particolare a quel matto in culo di Federico.
		Il rigraziamento più grande va alla mia ragazza e compagna di corso Federica, che mi ha sempre supportato e sopportato in questi anni.
}
 	\begin{abstract}
		In the last years, a new cyber-threat is spreading fast, known as \textit{Advanced Persistent Threat} (APT). An APT is a group that aims to penetrate and exflitrate information from high-valuable targets. They rely on malware and tools developed for the specific target they want to attack, usually governments, critical infrastructures or big organizations. Their goal is to break into the target's systems and stay undetected as long as possible. Meanwhile, they steal valuable information and try new ways to infect other target's computers. We propose a framework for malware prioritization to detect, among all the new malware samples that every day researchers find, the one that can be correlated to an APT group. The objective of this thesis is to correctly identify samples related to known APT. Those samples are then dispatched to analysts for further and manual inspection. To make this process as fast as possible we rely only on static features, without the need of executing the sample. We applied different techniques of feature selection to reduce the dataset dimensionality. We trained a \textit{RandomForestClassifier} with the selected features and we analyze its performances. In particular, we focus on precision because we do not want the analysts to waste time on false positives. Overall, we obtained great performances on both precision and accuracy, while runtime has been significantly reduced respect to the state-of-the-art.
	\end{abstract}
	\tableofcontents
	\mainmatter
	\input{introduction.tex}
	\input{relatedworks.tex}
	\input{tools.tex}
	\input{creation.tex}
	\input{classification.tex}
	\input{discussion.tex}
	\input{futureworks.tex}
	
	\backmatter
	\cleardoublepage
	\phantomsection % Give this command only if hyperref is loaded
	\addcontentsline{toc}{chapter}{\bibname}
	
	\bibliographystyle{ieeetr}

	
	% Here put the code for the bibliography. You can use BibTeX or
	% the BibLaTeX package or the simple environment thebibliography.
	\bibliography{reference}
\end{document}